---
title: "R Programing - Best Practices Part 2 (with answers)"
author: Mei Eisenbach
output:
    html_notebook: default
---
# R Programming Patterns
- Vectorization
- DRY (Do not Repeat yourself)
- Apply family of functions
- Functions and Lambda functions

## Vectorization
In R everything is a vector. That is to say, all variables are vectors.
```{r}
# a scalar
a <- 5
length(a)
```

```{r}
# vector
b <- c(1, 2, 3)
length(b)
```

```{r}
# string
s <- "foo"
length(s)
nchar(s)
```

```{r}
# data frame
a <- c(1, 2, 3)
b <- c(2, 3, 4)
df <- data.frame(a, b)
length(df)
```
A data frame is a vector of vectors. So the length is the number of columns.

### Vectorized Functions
Since all variables are vectors, built-in functions and functions from libraries are designed to handle them.

```{r}
# adding scalars
a <- 1
b <- 2
c <- a + b
c
```

```{r}
# adding vectors
a <- c(1, 2, 3)
b <- c(2, 3, 4)
c <- a + b
c
```

```{r}
# Sides of a right triangle
a <- c(3, 1, 1)
b <- c(4, 1, 2)

# TODO: Calculate and print the hypotenuse

```
Hint: $c = \sqrt(a^2 + b^2)$

## DRY (Do not Repeat yourself)
- "Every piece of knowledge must have a single, unambiguous, authoritative representation within a system"
- No copying and pasting code
- Programming principle formulated by Andy Hunt and Dave Thomas in their book The Pragmatic Programmer

```{r}
# create some fake data to work with
customer_count <- matrix(sample(c(1:100), 28), nrow = 4, ncol = 7)
customer_count
```
A matrix is collection of elements arranged in a two-dimensional rectangular layout. All the elements have the same type.

```{r}
# Bad:
# Get the means for each weekday
mon_mean <- mean(customer_count[,1])
tue_mean <- mean(customer_count[,2])
wed_mean <- mean(customer_count[,3])
thu_mean <- mean(customer_count[,5])
fri_mean <- mean(customer_count[,5])
sat_mean <- mean(customer_count[,6])
sun_mean <- mean(customer_count[,7])

weekday_means <- c(mon_mean, tue_mean, wed_mean, thu_mean, fri_mean, sat_mean, sun_mean)
weekday_means
```
Can you spot the bug?

## Apply() family
### apply()

apply() operates on arrays and matrices. 

Usage: apply(X, MARGIN, FUN, .)

where
X is an array or matrix
MARGIN indicates how the function is applied: 1 for rows, 2 for colums
FUN is the function applied to the data. Can be a user-defined function.
... are optional arguments for the function

```{r}
# Good:
# Get the means using apply on the columns
apply(customer_count, 2, mean)
```

```{r}
# TODO: Use the apply function to get totals per week


```

```{r}
# apply a function to each element in the matrix
apply(customer_count, c(1, 2), as.character)
```

```{r}
as.character(customer_count)
```
But why would I use apply when the as.character function is vectorized?

### lapply()
lapply() returns a list of the same length as X where each element in X has the function applied.

Usage: lapply(X, FUN, .)

where
X is an array or matrix
FUN is the function applied to the data
... are optional arguments

```{r}
data("iris")
head(iris)
```

```{r}
numeric_vars <- c(1:4)
lapply(iris[numeric_vars], mean)
```

```{r}
# TODO: get the standard deviations of each variable


```

## Functions
Function Definition

name.of.function <- function(argument1, argument2, ...) {
    statements
    return(something)
}

```{r}
get_mean <- function(x) {
  mean <- sum(x)/length(x)
  return(mean)
}

lapply(iris[numeric_vars], get_mean)

```


```{r}
# TODO: write a function that calculates the standard deviation


```

## Lambda functions
- Also known as anonymous functions.
- Should be used only for short, one line actions.
- Often used when the function is used only in one place.

```{r}
head(iris)

# Add one to each value
plus_one <- data.frame(lapply(iris[numeric_vars], (function(x) x+1)))
head(plus_one)
```

```{r}
# TODO: use a lambda function to multiple each value by 2
#   and then use it using lapply()



```

