---
title: "R Programing - Best Practices"
author: "Mei Eisenbach"
output:
    html_notebook: default
---
# Introduction
- Why this workshop
- What's covered
- What's not covered

# Workflow
## General tips
- Think about workflow first
- Use directories
- Use the features in your IDE

## Workspace
- Default behavior
- Saving the workspace is bad
- Clear Workspace: under the Session menu
- Code is what's important

## Projects
- Have their own workspace
- Working directory is set
- Can be created from existing directories

## Saving objects
- As an alternative to saving the workspace, save objects to files.
- saveRDS(), readRDS()

Demo
```{r message=FALSE, warning=FALSE}
library(randomForest)

data(iris)

# create train and test datasets
train_index <- sample(150, 0.80 * 150)
train <- iris[train_index,]
test <- iris[-train_index,]

# train random forest model
final_model <- randomForest(Species ~ ., data=iris)

#save the model to disk
saveRDS(final_model, "final_model.rds")
```

later...
```{r}
# load the model
loaded_model <- readRDS("final_model.rds")
loaded_model

# make predictions using loaded model on "new" data
preds <- predict(loaded_model, test)
```

## Github

# Style
## Naming

### Typical naming schemes
- Snake Case
- Period Separated
- Camel Case

### Naming tips
- Many workplaces will have their own standards
- Be consistent
- Strive for names that are concise and meaningful
- Variable names should be nouns and function names should be verbs

## Comments
- Comment your code!
- Comments should be informative

## Spacing
- Put a space before and after "->"
- Put a space before and after = when naming arguments in function calls. 
- Always put a space after a comma, never before, just like in regular English.

## Organize your code
- Use blank lines and comments to group sections of code.
- Place all dependencies at the beginning of the script.

# Styler
- Written in 2017 by Lorenz Walthert with Kirill Müller and Yihui Xie
- Once installed can be accessed from RStudio Add-ins
- Default style is tidyverse’s
- Customizable (https://lorenzwalthert.netlify.com/posts/stylerpost/)

```{r}
library(styler)
```

```{r}
ugly_code <- "a<-function( x){1+1}           "
style_text(ugly_code)
```

Other available commands:
- style_file() styles .R and/or .Rmd files.
- style_dir() styles all .R files in a directory.
- style_pkg() styles the source files of an R package.
- RStudio Addins for styling the active file, styling the current package and styling the highlighted code region.

```{r}
# Try the Addin:
#   - Select the line below.
#   - Go to the Addins menu and select "Style Selection"

a<-function( x){1+1}           

# To see that again, hit ctrl-Z to undo
```

```{r}
# Styler does not reformat variable names

variableName ="Camel Case"

```

```{r}
# Try typing some ugly code of your own and then fix it with styler



```

